Pangenome graph layout by Path-Guided Stochastic Gradient Descent

### 2023-09-14 manuscript/fig_sketches.tex

\begin{figure*}[!htb]
	\centering
		%\hline
		\includegraphics[width=\linewidth, trim=0cm 8cm 0cm 0cm, clip]{fig/sketches/PG-SGD.drawio.pdf}
	\caption{
		2D PG-SGD update operation sketches. \\
		\FIXME{ADD CAPTION.} \\ 
		\FIXME{PROVIDE SEVERAL NICE FIGURES SO WE CAN REARRANGE STUFF INTO SUBFIGURES.}
	}
	\label{fig:sketches}
\end{figure*}

>>>>

\begin{figure}[!htb]
	\centering
		%\hline
		\includegraphics[width=0.8\linewidth, trim=0cm 0cm 0cm 0cm, clip]{fig/sketches/FigureToHeaven.pdf}
	\caption{
		2D PG-SGD update operation sketches. \textbf{(a)} The path information of the graph. \textit{path1} and \textit{path2} both visit the same first node. Then their sequence diverges and they visit distinct nodes. \textbf{(b-e)} $v_i$/$v_j$ or $v_i$/$v_k$ is the current pair of nodes to update. $ld_{ij}$/$ld_{ik}$ is the current layout distance. $r,-r$ is the current size of the update. \textbf{(b)} Initial graph layout highlighting the future update of the two nodes of \textit{path1}. \textbf{(c)} The graph layout after the first update. The nodes appear longer now, because we updated at the end of the nodes. Highlighted is the future update of the two nodes of \textit{path2}. \textbf{(d)} The graph layout after the second update. Highlighted is the future update of the two nodes of \textit{path1}. \textbf{(e)} Final graph layout after three updates using the 2D PG-SGD.
	}
	\label{fig:sketches}
\end{figure}

### 2023-09-14 manuscript/main.tex

	Graph layout initialization can significantly influence the quality of the final layout.
	In the 1D implementation, nodes are placed in the same order as they appear in the input graph by default, although support for random layout initialization is also provided.
	In 2D, we offer several layout initialization techniques.
	One option is to position nodes in the first layout dimension according to their order in the input graph and to add either uniform or Gaussian noise in the second dimension.
	Another possibility is to arrange nodes along a Hilbert curve, which often favors the creation of planar final layouts.
    We additionally support fixing a set of node coordinates, such as those in a reference genome, in their order and position, which allows us to build 1D reference-focused views of complex pangenomes. An exapmle visualization is given in Figure~\ref{fig:1d_fig4}.
    \FIXME{Is this only for 1D? It would be \textit{amazing} if it were possible to do in 2D, and it should be easy. -EG} \\
    \FIXME{I am afraid this only works for 1D. Or what do you make of \url{https://github.com/pangenome/odgi/pull/491}? -SH}

>>>>

	Graph layout initialization can significantly influence the quality of the final layout.
	In the 1D implementation, nodes are placed in the same order as they appear in the input graph by default, although support for random layout initialization is also provided.
	In 2D, we offer several layout initialization techniques.
	One option is to position nodes in the first layout dimension according to their order in the input graph and to add either uniform or Gaussian noise in the second dimension.
	Another possibility is to arrange nodes along a Hilbert curve, which often favors the creation of planar final layouts.
    We additionally support fixing a set of node coordinates, such as those in a reference genome, in their order and position, which allows us to build reference-focused views of complex pangenomes. An example visualization is given in Figure~\ref{fig:1d_fig4}.
